Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-05-21 paper/paper.tex
tweak a little bit

\textbf{Abstracting Definitional Interpreters.}
Reynold's seminal paper \emph{Definitional interpreters for higher-order
  programming languages} \cite{reynolds1972definitional} shows that in
definitional interpreters, the defined language can inherit properties from the defining language.
With this insight, \citeauthor{darais2017abstracting} construct abstracting
definitional interpreters which automatically inherit the pushdown control-flow
property from its defining language
because the defined language simply uses the stack the
meta-language~\cite{darais2017abstracting}. \citeauthor{darais2017abstracting}'s abstracting
definitional interpreters work on direct-style programs, and is written in monadic style. One
of the advantages of monadic abstract interpreters is modular and composable, therefore
deploying different sensitivities or features is just applying a different monad.
Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.

This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work.
The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus,
instead of plain $\lambda$-calculus, although our work can also be easily adopted to handle a plain
$\lambda$-calculus.
But except of simply showing the final form of abstract definitional interpreters,
we reveal that refunctionalization plays an important role for inheriting the
stack from the metalanguage.
The target of the series of transformations in this paper is abstract definitional
interpreters, and we additionally demonstrate abstract interpreters written direct-style with
delimited controls. 
We use the delimited controls @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}.
The correspondence of delimited controls (as well as continuations) and monads
are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.

@@

We show the correspondence not only exists between concrete semantics artifacts but also exists between abstract semantics artifacts. An interesting open question would be whether there also exists a correspondence of static analyses formalized in different denotational style and operational style.

>>>>

\textbf{Abstracting Definitional Interpreters.}
Reynold's seminal paper \emph{Definitional interpreters for higher-order
  programming languages} \cite{reynolds1972definitional} shows that in
definitional interpreters, the defined language can inherit properties from the defining language.
With this insight, \citeauthor{darais2017abstracting} construct abstracting
definitional interpreters which automatically inherit the pushdown control-flow
property from its defining language because the defined language simply uses the stack of the
meta-language~\cite{darais2017abstracting}. \citeauthor{darais2017abstracting}'s abstracting
definitional interpreters work on direct-style programs, and is written in monadic style. One
of the advantages of monadic abstract interpreters is modular and composable, therefore
deploying different sensitivities or features is just applying a different monad.
Prior to that, \citeauthor{Sergey:2013:MAI:2491956.2491979} show a monadic
abstract interpreter for small-step semantics \cite{Sergey:2013:MAI:2491956.2491979}.

This paper is greatly inspired by~\citeauthor{darais2017abstracting}'s work.
The difference on the surface is that ours works for A-Normal Form $\lambda$-calculus,
instead of plain $\lambda$-calculus, although our work can also be easily adopted to handle a plain
$\lambda$-calculus.
But instead of directly showing the final form of abstract definitional interpreters,
we reveal that refunctionalization plays an important role for inheriting the
stack from the metalanguage.
The target of the series of transformations in this paper is abstract definitional
interpreters, and we additionally demonstrate abstract interpreters written direct-style with
delimited controls. 
We use the delimited control operators @shift@ and @reset@ that are implemented in Scala \cite{rompf2009implementing}.
The correspondence of delimited controls (as well as continuations) and monads
are well-known \cite{Danvy:1990:AC:91556.91622, wadler1992essence, danvy1992representing, moggi1991notions}.

@@

We show the functional correspondence not only exists between concrete semantics artifacts but also exists between abstract semantics artifacts. An interesting open question would be whether there also exists a correspondence of static analyses formalized in different denotational style and operational style.
